Neighborhoods and Segregation: An Information-Theoretic Lens

### 2016-07-03 slides.tex
fixes to short/long branching structure

\documentclass{beamer}
\usepackage{standalone}
\usepackage{import}
\usepackage{csquotes}
\usepackage{appendixnumberbeamer}
\usepackage{csvsimple}
\usepackage{bbding}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}

@@

\begin{document}
\maketitle
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}<handout:\included|beamer:\included>\frametitle{Table of contents}
	  \setbeamertemplate{section in toc}[sections numbered]
	  \tableofcontents[hideallsubsections]
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Neighborhoods and Diversity}
		Two questions for today: 
		\stepcounter{beamerpauses}
		\begin{enumerate}[<+ ->]
			\item What is \alert<2>{\textbf{urban diversity}}, and how should we measure it? 
			\item What is a \alert<3>{\textbf{neighborhood}}, and how can we distinguish neighborhoods in principled ways?
		\end{enumerate}
		\onslide<4->{
			\begin{block}{Central Claim}
				These two questions are deeply related, and we should approach both using \emph{information theory}.
			\end{block}}
			\note{This is a pretty damn testy note!}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Three Dimensions of Urban Diversity}

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Albuquerque: Low diversity, fairly even, high exposure}
		\begin{figure}
			\includegraphics[width=\textwidth]{external_figs/albuquerque_dot.png}
			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
		\end{figure}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Detroit: High diversity, highly uneven, low exposure}
		\begin{figure}
			\includegraphics[width=\textwidth]{external_figs/detroit_dot.png}
			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
		\end{figure}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Philadelphia: High diversity, highly uneven, high exposure}
		\begin{figure}
			\includegraphics[width=\textwidth]{external_figs/philly_dot.png}
			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
		\end{figure}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Diversity and Information Theory}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Why Information Theory?}
		Information theory quantifies concepts like uncertainty and predictability. Why use it for studying race? 
		\onslide<2->{
			\begin{block}{Core Insight}
				\begin{itemize}
					\item Complex systems allow \textbf{some} prediction...
					\item ...But good prediction is \textbf{hard}.
				\end{itemize}
			\end{block}
		}
		\onslide<3>{
			\textbf{Examples of prediction:}
			\begin{itemize}
				\item If you pick a random person from a city, is it hard to guess their race? ``Yes'' $\implies$ more diverse city. 
				\item Does it help if you tell me where they live? ``Yes'' $\implies$ more uneven city. 
			\end{itemize}
		}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Entropy as Global Diversity}
		Let $p(X,Y)$ be the joint distribution of location $X$ and race $Y$. The marginal distribution of race alone is $p(Y) = \sum_{x \in \mathcal{X}} p(x,Y)$. 
		\begin{block}{Entropy}
			\begin{equation*}
				H(Y) \triangleq - \mathbb{E}_Y[\log p(Y)] = -\sum_{y \in \mathcal{Y}} p(Y)\log p(Y)
			\end{equation*}
		\end{block}
		\onslide<2->{\begin{align*}
			H(Y) = 0 &\implies \text{ city is monoracial} \\
			H(Y) = \log \left|\mathcal{Y}\right| &\implies \text{ all races are represented equally}
		\end{align*}}
		\onslide<3->{\textbf{Entropy measures global diversity:} more entropy, more diversity.}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Mutual Information as Evenness}
		The mutual information quantifies the degree of dependence between two random variables.
		\begin{block}{Mutual Information}
			\begin{equation*}
			I(X,Y) \triangleq \sum_{x,y \in \mathcal{X} \times \mathcal{Y}}p(X,Y) \log \frac{p(X,Y)}{p(X)p(Y)}\;.
		\end{equation*}	
		\end{block}
		\onslide<2->{\begin{align*}
			I(X,Y) = 0 &\implies \text{ location and race are independent} \\
			I(X,Y) = \log \left|Y\right| &\implies \text{ location determines race}
		\end{align*}}
		\onslide<3->{\textbf{Mutual information measures evenness:} more information, the more uneven the spatial distribution of demographics.}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Local Information}
		Let $I_r(x_0)$ be local the mutual information between $X$ and $Y$, restricted to a small area of radius $r$ centered at $x_0$. Think of this as the \textbf{spatial variation of racial trends in a small area.} 
		\onslide<2->{
			\begin{block}{Theorem}
				Under appropriate regularity conditions, for small $r$,
				\begin{equation*}
					\frac{I_r(x_0)}{r^2} \cong \frac{1}{4} \text{trace } J_Y(x_0)\;, \tag{proof on request}
				\end{equation*}
				where $J_Y(x_0)$ is the Fisher information in $Y$ about $x_0$ \nocite{Amari2000}
			\end{block}
		}
		\onslide<3->{
			\alert{Significance}: the local information $I_r(x_0)$ is related to an intrinsic statistical measure $J_Y(x)$ \emph{that is independent of the resolution $r$}.
		}

@@

		\onslide<2->{
			\begin{block}{Mean Local Information}
				\begin{equation*}
					J(X,Y) \triangleq \mathbb{E}_X[\text{trace } J_Y(X)]\;.
				\end{equation*}
			\end{block}
		}
		\onslide<3->{
			\begin{align*}
				J(X,Y) = 0 &\implies \text{No neighborhoods, low exposure} \\
				J(X,Y) \text{ large} &\implies \text{Lots of neighborhoods, high exposure}
			\end{align*}
		}
		\onslide<4->{
			This is one way of ``spatializing'' the mutual information $I(X,Y)$; for another see \cite{Roberto2015a,Roberto2015}. 
		}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Back to the checkerboard}
		\begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/checkerboard.png}
	    \end{figure}
	    \centering
	    \begin{tabular}{l c c c}
			  \textbf{City} & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
			  \hline			
			  (a) & \alert{0.0} & 0.0 & 0.0\\
			  (b) & \alert{0.7} & \alert{0.0} & 0.0\\
			  (c) & 0.7 & \alert{0.7} & \alert{0.6}\\
			  (d) & 0.7 & 0.7 & \alert{2.7}\\
		\end{tabular}

@@

    	\textit{Data accessed from the American Community Survey of the U.S. Census \cite{CensusRace}.
	}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Visualizing Diversity Profiles in Major Cities}
		\centering
		\begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/mutual_fisher.png}
	    \end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Urban Density and the Compression of Social Space}
		\centering
		\begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/density_fisher.png}
	    \end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Information-Theoretic Identification of Neighborhoods}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{How can we use this?}
		$H(Y)$, $I(X,Y)$, and $J(X,Y)$ measure diversity, evenness, and spatial exposure.

@@

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[t]\frametitle{Key Property of Mutual Information}
	    Suppose we cluster the locations $X$ into groups with labels $C$. 
	    \begin{block}{\only<1-2>{``Additive Organizational Decomposability'' (Chain Rule)} \only<3->{Recipe for Neighborhood Identification} }
		    \begin{equation*}
		    	\only<1-2>{I(X,Y) = \onslide<1-2>{I(C,Y)} + \onslide<1-2>{I(X,Y|C)}\;.}
		    	\only<3->{I(X,Y) = \underbrace{I(C,Y)}_\text{maximize this} + \overbrace{I(X,Y|C)}^\text{minimize this}\;.}
		    \end{equation*}
	    \end{block}

@@

	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}{Open Directions}
		Let's learn about...
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[standout]
		Thank you!

>>>>

\documentclass{beamer}
\usepackage{standalone}
\usepackage{import}
\usepackage{csquotes}
\usepackage{appendixnumberbeamer}
\usepackage{csvsimple}
\usepackage{bbding}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{etoolbox}

@@

\begin{document}

% \def\short{1}

\newif\iflong
\ifdefined\short
	\longfalse
\else
	\longtrue
\fi

\maketitle
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
		\begin{frame}\frametitle{Table of contents}
		  \setbeamertemplate{section in toc}[sections numbered]
		  \tableofcontents[hideallsubsections]
		\end{frame}
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Introduction}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\begin{frame}\frametitle{Neighborhoods and Diversity}
		Two questions for today: 
		\stepcounter{beamerpauses}
		\begin{enumerate}[<+ ->]
			\item What is \alert<2>{\textbf{urban diversity}}, and how should we measure it? 
			\item What is a \alert<3>{\textbf{neighborhood}}, and how can we distinguish neighborhoods in principled ways?
		\end{enumerate}
		\onslide<4->{
			\begin{block}{Central Claim}
				These two questions are deeply related, and we should approach both using \emph{information theory}.
			\end{block}}
			\note{This is a pretty damn testy note!}
	\end{frame}

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Three Dimensions of Urban Diversity}

@@

	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
	\begin{frame}\frametitle{Albuquerque: Low diversity, fairly even, high exposure}
		\begin{figure}
			\includegraphics[width=\textwidth]{external_figs/albuquerque_dot.png}
			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
		\end{figure}
	\end{frame}
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
	\begin{frame}\frametitle{Detroit: High diversity, highly uneven, low exposure}
		\begin{figure}
			\includegraphics[width=\textwidth]{external_figs/detroit_dot.png}
			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
		\end{figure}
	\end{frame}
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
	\begin{frame}\frametitle{Philadelphia: High diversity, highly uneven, high exposure}
		\begin{figure}
			\includegraphics[width=\textwidth]{external_figs/philly_dot.png}
			\caption{\textit{Image Copyright, 2013, Weldon Cooper Center for Public Service, Rector and Visitors of the University of Virginia (Dustin A. Cable, creator)}}
		\end{figure}
	\end{frame}
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Diversity and Information Theory}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
		\begin{frame}\frametitle{Why Information Theory?}
			Information theory quantifies concepts like uncertainty and predictability. Why use it for studying race? 
			\only<2->{
				\begin{block}{Core Insight}
					\begin{itemize}
						\item Complex systems allow \textbf{some} prediction...
						\item ...But good prediction is \textbf{hard}.
					\end{itemize}
				\end{block}
			}
			\onslide<3>{
				\textbf{Examples of prediction:}
				\begin{itemize}
					\item If you pick a random person from a city, is it hard to guess their race? ``Yes'' $\implies$ more diverse city. 
					\item Does it help if you tell me where they live? ``Yes'' $\implies$ more uneven city. 
				\end{itemize}
			}
		\end{frame}

	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	\iflong
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Entropy as Global Diversity}
		Let $p(X,Y)$ be the joint distribution of location $X$ and race $Y$. The marginal distribution of race alone is $p(Y) = \sum_{x \in \mathcal{X}} p(x,Y)$. 
		\begin{block}{Entropy}
			\begin{equation*}
				H(Y) \triangleq - \mathbb{E}_Y[\log p(Y)] = -\sum_{y \in \mathcal{Y}} p(Y)\log p(Y)
			\end{equation*}
		\end{block}
		\onslide<2->{\begin{align*}
			H(Y) = 0 &\implies \text{ city is monoracial} \\
			H(Y) = \log \left|\mathcal{Y}\right| &\implies \text{ all races are represented equally}
		\end{align*}}
		\onslide<3->{\textbf{Entropy measures global diversity:} more entropy, more diversity.}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Mutual Information as Evenness}
		The mutual information quantifies the degree of dependence between two random variables.
		\begin{block}{Mutual Information}
			\begin{equation*}
			I(X,Y) \triangleq \sum_{x,y \in \mathcal{X} \times \mathcal{Y}}p(X,Y) \log \frac{p(X,Y)}{p(X)p(Y)}\;.
		\end{equation*}	
		\end{block}
		\onslide<2->{\begin{align*}
			I(X,Y) = 0 &\implies \text{ location and race are independent} \\
			I(X,Y) = \log \left|Y\right| &\implies \text{ location determines race}
		\end{align*}}
		\onslide<3->{\textbf{Mutual information measures evenness:} more information, the more uneven the spatial distribution of demographics.}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Local Information}
		Let $I_r(x_0)$ be local the mutual information between $X$ and $Y$, restricted to a small area of radius $r$ centered at $x_0$. Think of this as the \textbf{spatial variation of racial trends in a small area.} 
		\onslide<2->{
			\begin{block}{Theorem}
				Under appropriate regularity conditions, for small $r$,
				\begin{equation*}
					\frac{I_r(x_0)}{r^2} \cong \frac{1}{4} \text{trace } J_Y(x_0)\;, \tag{proof on request}
				\end{equation*}
				where $J_Y(x_0)$ is the Fisher information in $Y$ about $x_0$ \nocite{Amari2000}
			\end{block}
		}
		\onslide<3->{
			\alert{Significance}: the local information $I_r(x_0)$ is related to an intrinsic statistical measure $J_Y(x)$ \emph{that is independent of the resolution $r$}.
		}

@@

		\onslide<2->{
			\begin{block}{Mean Local Information}
				\begin{equation*}
					J(X,Y) \triangleq \mathbb{E}_X[\text{trace } J_Y(X)]\;.
				\end{equation*}
			\end{block}
		}
		\onslide<3->{
			\begin{align*}
				J(X,Y) = 0 &\implies \text{No neighborhoods, low exposure} \\
				J(X,Y) \text{ large} &\implies \text{Lots of neighborhoods, high exposure}
			\end{align*}
		}
		\onslide<4->{
			This is one way of ``spatializing'' the mutual information $I(X,Y)$; for another see \cite{Roberto2015a,Roberto2015}. 
		}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\fi
	\iflong
		\longfalse
	\else
		\longtrue
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
	\begin{frame}\frametitle{Summary: Information Theory and Diversity}
		\begin{block}{Entropy Measures Diversity}
			$H(Y) \triangleq - \mathbb{E}_Y[\log p(Y)]$
		\end{block}
		\onslide<2->{
			\begin{block}{Mutual Information Measures Evenness}	
				$I(X,Y) \triangleq \mathbb{E}_{X,Y}\left[\log \frac{p(X,Y)}{p(X)p(Y)}\right]$
			\end{block}
		}
		\onslide<3->{
			\begin{block}{Mean Local Information Measures Exposure}
				$J(X,Y) \triangleq \mathbb{E}_X [\text{trace } J_Y(X)]$
			\end{block}	
		}	
	\end{frame}
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
		\longfalse
	\else
		\longtrue
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Back to the checkerboard}
		\begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/checkerboard.png}
	    \end{figure}
	    \centering
	    \begin{tabular}{l c c c}
			  \textbf{City} & $H(Y)$ & $I(X,Y)$ & $J(X,Y)$ \\
			  \hline			
			  (a) & \alert{0.0} & 0.0 & 0.0\\
			  (b) & \alert{0.7} & \alert{0.0} & 0.0\\
			  (c) & 0.7 & \alert{0.7} & \alert{0.6}\\
			  (d) & 0.7 & 0.7 & \alert{2.7}\\
		\end{tabular}

@@

    	\textit{Data accessed from the American Community Survey of the U.S. Census \cite{CensusRace}.
	}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}\frametitle{Visualizing Diversity Profiles in Major Cities}
		\centering
		\begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/mutual_fisher.png}
	    \end{figure}
	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
	\begin{frame}\frametitle{Urban Density and the Compression of Social Space}
		\centering
		\begin{figure}
	    	\centering
	    	\includegraphics[width=\textwidth]{figs/density_fisher.png}
	    \end{figure}
	\end{frame}
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Information-Theoretic Identification of Neighborhoods}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
	\begin{frame}\frametitle{How can we use this?}
		$H(Y)$, $I(X,Y)$, and $J(X,Y)$ measure diversity, evenness, and spatial exposure.

@@

	\end{frame}
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[t]\frametitle{Key Property of Mutual Information}
	    Suppose we cluster the locations $X$ into groups with labels $C$. 
	    \begin{block}{\only<1-2>{``Additive Organizational Decomposability'' (Chain Rule)} \only<3->{Recipe for Neighborhood Identification} }
		    \begin{equation*}
		    	\only<1-2>{I(X,Y) = \onslide<1-2>{I(C,Y)} + \onslide<1-2>{I(X,Y|C)}\;.}
		    	\only<3->{I(X,Y) = \underbrace{I(C,Y)}_\text{maximize this} + \overbrace{I(X,Y|C)}^\text{minimize this}\;.}
		    \end{equation*}
	    \end{block}

@@

	\end{frame}
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\iflong
	\begin{frame}{Open Directions}
		Let's learn about...
	\end{frame}
	\fi
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
	\begin{frame}[standout]
		Thank you!
